Contextual Combinatorial Cascading Bandits

### 2015-12-04 contextual combinatorial cascading bandits.tex

We formulate the problem of {\em contextual combinatorial cascading bandit} as follows. Suppose we have a finite set of $L$ ground items,  $E=\{1,...,L\}$, also referred to as {\em base arms}. 
Let $\prod^k=\{(a_1,...,a_k): a_1,...,a_k \in E, a_i \neq a_j \text{ for any } i \neq j\}$ be the set of all $k$-tuples of distinct items from $E$. We also refer to each of such tuples as an {\em action}.
Let $\cS \subseteq \cup_{k=1}^K \prod^{k}$ be the set of feasible actions with length no more than $K$. As a convention, we use boldface symbols to represent all random variables.

At time $t$, feature vectors $x_{t,a} \in \RR^{d \times 1}$ with $\norm{x_{t,a}}_2 \leq 1$ for every base arm $a \in E$ are revealed to the learning agent; the feature vectors combine both the information of the user and the corresponding base arm. Then the learning agent recommends a feasible action $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ to the user. 
\wei{Why using $\ba_{1}^t$, not $a_1^t$?} \shuai{I think $a_1^t$ is a random variable because $\bA_t$ is random. Is it making sense?}
Starting from the first item, the user checks the items one by one. The checking process stops if the user clicks on one item or has checked all items without clicking anyone. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ as the {\em weight} of base arm $a$ at time $t$ to indicate whether the user has clicked on $a$ or not.

Let $\cH_t$ denote all history before the learning agent chooses action at time $t$; $\cH_t$ contains all information and observations at time $s < t$ and context information at time $t$. Given history $\cH_t$, we assume $\bw_{t,a}$'s are mutually independent and satisfy
\wei{Is this given $x_{t,a}$ accurate? $x_{t,a}$  is not a probabilistic event. Also, are we assuming $\bw_{t,a}$'s are
		also independent of all other events, such as the event before time $t$? We should make it clear.
Perhaps we should put the history definition here?}
\begin{equation}
\label{eq:expectation}
\EE[\bw_{t,a} | \cH_t] = \theta_{\ast}^{\top} x_{t,a},
\end{equation}
where $\theta_{\ast}$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_{\ast}}_2 \leq 1$ and $0 \leq \theta_{\ast}^{\top} x_{t,a} \leq 1$ for all $t, a$. 
We define random variable $\bO_t$ as the number of observed base arms in $\bA_t$, that is,
	for some $k=1,2,\ldots, \abs{\bA_t}$, 
$$
\bO_t = k, \text{ if } \bw_t(\ba_j^t)=0, \forall\, j < k \text{ and } \bw_t(\ba_k^t) = 1,
$$
or 
$$
\bO_t = \abs{\bA_t}, \text{ if }\bw_t(\ba_j^t) = 0, ~~ \forall\, j \leq \abs{\bA_t}.
$$
At the end of every time $t$, the agent observes the first $\bO_t$ items of $\bA_t$. We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$.

@@

Because $\cH_t$ denotes all history before time $t$ and context information of time $t$, $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s), k \in[\bO_s], s<t, x_t \}$ and define $\EE_t[\cdot] = \EE[\cdot | \cH_t]$. Here we abuse the notation $\abs{\bA_s}$ to denote the number of base arms in the action $\bA_s$. By Equation \eqref{eq:expectation}, we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k x_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a positive invertible matrix. In the following, we will mainly use notations for disjunctive objective when the two cases are similar.

@@

\begin{theorem}[\cite{abbasi2011improved}]
\label{thm:theta_estimate}
Let 
$$
\beta_{t}(\delta) = \sqrt{\log(\det(\bV_{t})) + 2 \log(\frac{1}{\delta})} + \sqrt{\lambda}.
$$
Then for any $\delta > 0$, with probability at least $1 - \delta$, for all $t > 0$, we have
\begin{equation}
\label{eq:estimateTheta}
\norm{\hat{\theta}_t - \theta_{\ast}}_{\bV_{t}} \leq \beta_{t}(\delta).
\end{equation}
\end{theorem}

>>>>

We formulate the problem of {\em contextual combinatorial cascading bandit} as follows. Suppose we have a finite set of $L$ ground items,  $E=\{1,...,L\}$, also referred to as {\em base arms}. 
Let $\prod^k=\{(a_1,...,a_k): a_1,...,a_k \in E, a_i \neq a_j \text{ for any } i \neq j\}$ be the set of all $k$-tuples of distinct items from $E$. We also refer to each of such tuples as an {\em action}.
Let $\cS \subseteq \cup_{k=1}^K \prod^{k}$ be the set of feasible actions with length no more than $K$. 
As a convention, we always use boldface symbols to represent random variables.

At time $t$, feature vectors $x_{t,a} \in \RR^{d \times 1}$ with $\norm{x_{t,a}}_2 \leq 1$ for every base arm $a \in E$ are revealed to the learning agent; the feature vectors combine both the information of the user and the corresponding base arm. Then the learning agent recommends a feasible action $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ to the user. 
Starting from the first item, the user checks the items one by one. The checking process stops if the user clicks on one item or has checked all items without clicking anyone. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ as the {\em weight} of base arm $a$ at time $t$ to indicate whether the user has clicked on $a$ or not.

Let $\cH_t$ denote the history before the learning agent chooses action at time $t$.
Thus $\cH_t$ contains all information and observations at time $s < t$ and context information at time $t$. Given history $\cH_t$, we assume $\bw_{t,a}$'s are mutually independent and satisfy
\begin{equation}
\label{eq:expectation}
\EE[\bw_{t,a} | \cH_t] = \theta_{\ast}^{\top} x_{t,a},
\end{equation}
where $\theta_{\ast}$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_{\ast}}_2 \leq 1$ and $0 \leq \theta_{\ast}^{\top} x_{t,a} \leq 1$ for all $t, a$. 
We define random variable $\bO_t$ as the number of observed base arms in $\bA_t$, that is,
	for some $k=1,2,\ldots, \abs{\bA_t}$, 
$$
\bO_t = k, \text{ if } \bw_t(\ba_j^t)=0, \forall\, j < k \text{ and } \bw_t(\ba_k^t) = 1,
$$
or 
$$
\bO_t = \abs{\bA_t}, \text{ if }\bw_t(\ba_j^t) = 0, ~~ \forall\, j \leq \abs{\bA_t}.
$$
At the end of every time $t$, the agent observes the first $\bO_t$ items of $\bA_t$. We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$. 
Thus, $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s),x_t 
	\mid  k \in[\bO_s], 1 \le s<t \}$.

@@

At time $t$, we denote $\EE_t[\cdot] = \EE[\cdot | \cH_t]$, where $\cH_t$ is the history a time $t$ before the learning agent
	chooses her action. 
By Equation \eqref{eq:expectation}, we have 
$$
\EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
\{(\gamma_k x_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
\hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
\bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a positive invertible matrix. 
%In the following, we will mainly use notations for disjunctive objective when the two cases are similar.

@@

\begin{theorem}[\cite{abbasi2011improved}]
\label{thm:theta_estimate}
Let 
$$
\beta_{t}(\delta) = \sqrt{\log(\det(\bV_{t})) + 2 \log\left(\frac{1}{\delta}\right)} + \sqrt{\lambda}.
$$
Then for any $\delta > 0$, with probability at least $1 - \delta$, for all $t > 0$, we have
\begin{equation}
\label{eq:estimateTheta}
\norm{\hat{\theta}_t - \theta_{\ast}}_{\bV_{t}} \leq \beta_{t}(\delta).
\end{equation}
\end{theorem}
